MG7: Configurable and scalable 16S metagenomics data analysis -- new methods optimized for massive cloud computing

### 2015-09-18 docs/paper/out/paper.tex
make minor wording corrections on Discussion -- paper #35

On one hand, specification of the input data allows us to restrict its
type and force users to be conscious about what they pass as an input.
On the other hand specification of the output data helps to build a
workflow as a \emph{composition} of several parts: we can ensure on the
Scala code type level that the output of one component fits as an input
of the next component. This can be crucial, as often the way the
analysis works depends a lot on the particular structure of the data.
For instance, in the MG7 workflow, using BLAST eDSL, we can describe
exactly which format will the output of the BLAST step have, which
information it will include, and then in the next step we can reuse this
description to easily parse BLAST output and retrieve the part of the
information needed for the taxonomy assignment analysis. Having data
structure described statically as Scala code allows us to be sure that
we won't have parsing problems or other issues with incompatible data
passed between components of the workflow.

All this doesn't compromise flexibility of the way user works with data
in MG7. On the contrary, having static data declarations as a part of
the configuration allows user to reuse component of analysis and modify
it easily according to particular needs. Besides that, an important
advantage of the type-level control is the additional insurance from
unsuccessful analysis launches, which may lead to the lost of time and
as a consequence finance spent on the cloud resources.

@@

Besides the bioinformatics tools like BLAST and FLASH, \emph{Statika}
bundles are used for wrapping data dependencies and all inner components
of the system that require cloud deployment. In particular, all
components of \emph{Loquat} are bundles, which allows user to define
which components are needed for the parallel processing on each
computation unit in an expressive way, declaring them as bundle
dependencies of the loquat ``worker'' bundle. This modularization is
also important for the matter of making components of the system easily
reusable for different projects and liberating user from most of the
tasks related to their deployment.

@@

The MG7 workflow consists of certain steps, each of which perform some
work in parallel, using the cloud infrastructure managed by
\emph{Loquat}. It is important to notice the horizontal scalability of
this approach. Irrespectively of how much data is needed to be
proccessed, MG7 will easily handle it, by splitting data on chunks and
performing the analysis on multiple computaion units. The Amazon Elastic
Compute Cloud (EC2) provides a transparent way of managing computation
infrastructure, called autoscaling groups. User can easily set up MG7
configuration parameters, adjusting the amount of EC2 instances they
want to occupy for each task and their type. When using Amazon Web
Services for parallel computations, it is easy to scale the computation
resources, because the you pay for the time, not the amount of
computation units.

@@

\begin{itemize}
\tightlist
\item
  EC2 (Elastic Compute Cloud) autoscaling groups for launching and
  managing computation units
\item
  S3 (Simple Storage Service) for storing input and output data
\item
  SQS (Simple Queue Service) for communication between different
  components of the system
\item
  SNS (Simple Notification Service) for notifying user about the
  progress
\end{itemize}

>>>>

On one hand, specification of the input data allows us to restrict its
type and force users to be conscious about what they pass as an input.
On the other hand specification of the output data helps to build a
workflow as a \emph{composition} of several parts: we can ensure on the
Scala code type level that the output of one component fits as an input
for the next component. This is crucial, as obviously the way a data
analysis task works depends a lot on the particular structure of the
data. For instance, in the MG7 workflow, using BLAST eDSL, we can
precisely describe which format will have the output of the BLAST step,
which information it will include, and then in the next step we can
reuse this description to parse BLAST output and retrieve the part of
the information needed for the taxonomy assignment analysis. Having data
structure described statically as Scala code allows us to be sure that
we won't have parsing problems or other issues with incompatible data
passed between workflow components.

All this does not compromise flexibility in how the user works with data
in MG7: having static data declarations as a part of the configuration
allows the user to reuse analysis components, or modify them according
to particular needs. Besides that, an important advantage of the
type-level control is the added protection from the execution (and
deployment) of a wrongly configured analysis task, which may lead to
significant costs in both time and money.

@@

Besides the bioinformatics tools like BLAST and FLASH, \emph{Statika}
bundles are used for wrapping data dependencies and all inner components
of the system that require cloud deployment. In particular, all
components of \emph{Loquat} are bundles; the user can then define which
components are needed for the parallel processing on each computation
unit in an expressive way, declaring them as bundle dependencies of the
loquat ``worker'' bundle. This modularization is also important for the
matter of making components of the system reusable for different
projects and liberating the user from most of the tasks related to their
deployment.

@@

The MG7 workflow consists of certain steps, each of which perform some
work in parallel, using the cloud infrastructure managed by
\emph{Loquat}. It is important to notice the horizontal scalability of
this approach. Irrespectively of how much data needs to be processed,
MG7 will easily handle it, by splitting data on chunks and performing
the analysis on multiple computation units. The Amazon Elastic Compute
Cloud (EC2) service provides a transparent way of managing computation
infrastructure, called autoscaling groups. The User can set MG7
configuration parameters, adjusting for each task the amount and
hardware characteristics of the EC2 instances they want to occupy for
it. When using Amazon Web Services for parallel computations, it is easy
to scale the computation resources, because the you pay for the time,
not the amount of computation units.

@@

\begin{itemize}
\tightlist
\item
  \href{https://aws.amazon.com/ec2}{EC2} (Elastic Compute Cloud)
  autoscaling groups for launching and managing computation units
\item
  \href{https://aws.amazon.com/s3}{S3} (Simple Storage Service) for
  storing input and output data
\item
  \href{https://aws.amazon.com/sqs}{SQS} (Simple Queue Service) for
  communication between different components of the system
\item
  \href{https://aws.amazon.com/sns}{SNS} (Simple Notification Service)
  for e-mail notifications
\end{itemize}
